Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-11-13 paper/1_Initial/manuscript.qmd
minor updates

This pattern may help clarify inconsistencies in the literature linking interoception and emotion. Reviews and meta-analyses report that performance-based measures of interoceptive accuracy, particularly in cardioception, often show weak or absent correlations with depression and anxiety [@jenkinson2024interoception; @banellis2025interoceptive; @adams2022association]. In contrast, self-reported interoceptive measures typically show stronger, often negative, associations with internalizing traits [e.g., @brand2022; @lin2023; @clemente2024relationship]. These discrepancies likely reflect both conceptual distinctions between perceptual accuracy and metacognitive self-belief [@khalsa2018; @garfinkel2015knowing; @suksasilp2022towards] and variation in the bodily domains assessed. The present results extend this framework by demonstrating that associations between interoceptive self-beliefs and affective traits are concentrated in psychologically salient visceral domains. Thus, the field’s heavy reliance on cardioceptive tasks may have obscured more meaningful relationships in other interoceptive modalities.

<!-- Psychopathology -->
Beyond mood, similar though weaker patterns emerged for autistic traits, schizotypy, and maladaptive personality dimensions. These traits tended to correlate negatively with self-perceived interoceptive accuracy, particularly within visceral and musculoskeletal domains, while showing little or no association in other areas. This selective pattern suggests that diminished confidence in bodily perception may contribute to specific social–emotional and personality features, rather than reflecting a general interoceptive deficit. These results align with evidence that interoceptive self-beliefs, more than objective accuracy, track variation in emotional awareness, social connectedness, and psychopathology [@torregrossa2022interoceptive; @williams2023characterizing; @brand2022].

>>>>

This pattern may help clarify inconsistencies in the literature linking interoception and emotion. Reviews and meta-analyses report that performance-based measures of interoceptive accuracy, usually cardioception, often show weak or absent correlations with depression and anxiety [@jenkinson2024interoception; @banellis2025interoceptive; @adams2022association]. In contrast, self-reported interoceptive measures, which contain items relevant to multiple bodily domains, typically show stronger, often negative, associations with internalizing traits [e.g., @brand2022; @lin2023; @clemente2024relationship]. These discrepancies likely reflect both conceptual distinctions between perceptual accuracy and metacognitive self-belief [@khalsa2018; @garfinkel2015knowing; @suksasilp2022towards] and variation in the bodily domains assessed. The present results extend this framework by demonstrating that associations between interoceptive self-beliefs and affective traits are concentrated in psychologically salient visceral domains. Thus, the field’s heavy reliance on cardioceptive tasks may have obscured more meaningful relationships in other interoceptive modalities.

<!-- Psychopathology -->
Beyond mood, similar though weaker patterns emerged for autistic traits, schizotypy, and maladaptive personality dimensions. These traits tended to correlate negatively with self-perceived interoceptive accuracy, particularly within visceral and musculoskeletal domains, while showing little or no association in other areas. This selective pattern suggests that diminished beliefs in bodily perception may contribute to specific social–emotional and personality features, rather than reflecting a general interoceptive deficit. These results align with evidence that interoceptive self-beliefs, more than objective accuracy, track variation in emotional awareness, social connectedness, and psychopathology [@torregrossa2022interoceptive; @williams2023characterizing; @brand2022].
